---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


Firstly I load the dataset from the zip files. 
```{r echo = TRUE}
library(readr)
data <- read_csv("activity.zip")
```

Secondly I wanted to show some first elements of this dataset to get to know something about them. Of course I used the 'summary' and 'str' function to see some basics statistics.
```{r echo = TRUE}
summary(data)
head(data)
str(data)
dim(data)
```

Now I know that this dataset have a lot of NA rows. My goal is to recognize why this value is so common. Below I show how many this NA value are. 

```{r echo = TRUE}
Percentage_NA <- length(which(is.na(data$steps)))/length(data$steps)
print(Percentage_NA)

```
The reason of so many NA values in this dataset is very simple - in that days this device didn't work or was turned off. 


## What is mean total number of steps taken per day?

```{r echo = TRUE, }
library(dplyr)

sum_perday <- data %>%
  group_by(date) %>%
  summarise(sum_perday = sum(steps, na.rm = TRUE))

mean_steps <- summarize(sum_perday, mean_perday = round(mean(sum_perday, na.rm = TRUE), 2))
```
Answer: Mean total number of steps taken per day is equal to `r mean_steps`.

## What is the average daily activity pattern?



## Imputing missing values

```{r echo = TRUE}
data_imput <- filter(data, !is.na(steps))
head(data_imput)
```

## Are there differences in activity patterns between weekdays and weekends?
